### 2013-06-07 Tex/chapters/model.tex
Besseres Erwartungswert Zeichen

\subsection{Model}
    The used model is a modified 2D Ising model where the \(N\)sites are
    displaced nodes of a square lattice of edgelength \(L\) with
    periodic boundary conditions. The displacement is randomly gauss
    distributed with the standard deviation \(\sigma\). This \(\sigma\)
    is also called \emph{disturbance paramter} in the following.
    The hamiltonian of the Ising model is
    \(\hat{H} = \sum_{<i,j>}J_{ij}s_{i}s_{j}\)
    where \(<i,j>\) refers to by edges connected nodes.
    The edges are constructed according to the in section
    \ref{sec:graphtypes} defined rules. So that the lattice resembles a
    proximity graph. Each node \(i\) has a spin \(s_i = \pm 1\). Each
    edge has a weight \(J_{ij} = \exp (\alpha (1-d_{ij}))\) with the
    distance \(d_{ij}\) between the nodes \(i\) and \(j\). \(J\) is
    called \emph{interaction}.\\
    For \(\sigma = 0\) this is the standard Ising model with \(J = 1\),
    for which exists an analytical solution \cite{Onsager1944}.\\

    For evaluation the Monte Carlo simulation is is run till the system
    is equilibrated after \(t_eq\) sweeps. Then the simulation continues
    and the magnetisation per site \(m\) and energy of the system \(E\)
    are calculated and saved for every \(2\tau\) sweeps. Where \(\tau\)
    denotes the \emph{autocorrelation time} after which two states of
    the system are not correlated anymore.
    For every observable \(O\) the expected value \(<O>\) is averaged
    over different randomly generated lattices \(\overline{<O>}\). The
    errors are estimated by bootstrapping.

\subsection{Critical Temperature $T_c$ and Critical Exponents}
    Because the searched for property -- the critical temperature \(T_c\)
    -- is only defined for infinite systems, hence every computer
    simulation will show some \emph{finte size effects}.
    [vielleicht Bild von intensiver Größe?]
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{Newman1998} which also yield the critical
    exponents. The critical exponents should not be influenced by the
    disturbance paramter \(\sigma\) [citation needed]. So they will be
    used for consistency cross checking and comparision with the known
    exact values.\\
    [weiter ausarbeiten] [sagen, dass man für zwei/drei sigma
    die exponenten reproduziert hat].\\
    Though if one is just interested in the critical Temperature an
    easier approach is to find the intersections of the binder cumulants
    \(g = (3-\frac{<m^4>}{2<m^2>^2}\) \cite{Binder1981} of different
    system sizes \(L=\sqrt N\).\\

>>>>

\subsection{Model}
    The used model is a modified 2D Ising model where the \(N\)sites are
    displaced nodes of a square lattice of edgelength \(L\) with
    periodic boundary conditions. The displacement is randomly gauss
    distributed with the standard deviation \(\sigma\). This \(\sigma\)
    is also called \emph{disturbance paramter} in the following.
    The hamiltonian of the Ising model is
    \(\hat{H} = \sum_{\avg{i,j}}J_{ij}s_{i}s_{j}\)
    where \(\avg{i,j}\) refers to by edges connected nodes.
    The edges are constructed according to the in section
    \ref{sec:graphtypes} defined rules. So that the lattice resembles a
    proximity graph. Each node \(i\) has a spin \(s_i = \pm 1\). Each
    edge has a weight \(J_{ij} = \exp (\alpha (1-d_{ij}))\) with the
    distance \(d_{ij}\) between the nodes \(i\) and \(j\). \(J\) is
    called \emph{interaction}.\\
    For \(\sigma = 0\) this is the standard Ising model with \(J = 1\),
    for which exists an analytical solution \cite{Onsager1944}.\\

    For evaluation the Monte Carlo simulation is is run till the system
    is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
    and the magnetisation per site \(m\) and energy of the system \(E\)
    are calculated and saved for every \(2\tau\) sweeps. Where \(\tau\)
    denotes the \emph{autocorrelation time} after which two states of
    the system are not correlated anymore.
    For every observable \(O\) the expected value \(\avg{O}\) is averaged
    over different randomly generated lattices \(\overline{\avg{O}}\). The
    errors are estimated by bootstrapping.

\subsection{Critical Temperature $T_c$ and Critical Exponents}
    Because the searched for property -- the critical temperature \(T_c\)
    -- is only defined for infinite systems, hence every computer
    simulation will show some \emph{finte size effects}.
    [vielleicht Bild von intensiver Größe?]
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{Newman1998} which also yield the critical
    exponents. The critical exponents should not be influenced by the
    disturbance paramter \(\sigma\) [citation needed]. So they will be
    used for consistency cross checking and comparision with the known
    exact values \cite[S. 59]{Pelissetto2002}.\\
    %~ [theorie mit dem Collaps]
    %~ [Erklärung autoscale]
    %~ [Am besten hier direkt die ganze Auswertung rein]
    %~ [weiter ausarbeiten] [sagen, dass man für zwei/drei sigma
    die exponenten reproduziert hat].\\
    Though if one is just interested in the critical Temperature an
    easier approach is to find the intersections of the binder cumulants
    \(g = \brac{3-\frac{\avg{m^4}}{2\avg{m^2}^2}}\) \cite{Binder1981} of different
    system sizes \(L=\sqrt N\).
    Therefore a polynom of degree 4 is fitted to \(g\) in vicinity of
    the intersection. %[vgl. Grafik (villeicht)]
    The arithmetic average of the intersections are plotted [hier schon?]
    %~ [Und den rest der Auswertung]
    \\
